---
title: "Dealing with missing values"
output:
  html_notebook: default
  html_document: default
---

Titanic data set. This is a data set that records various attributes of passengers on the Titanic, including who survived and who didn’t.

```{r Library Files, message=FALSE, warning=FALSE, include=FALSE}
install.packages("stringdist")
library("stringdist")
install.packages("dplyr")
library("dplyr")
install.packages("tidyr")
library("tidyr")
install.packages("dummies")
library("dummies")
```

Save the data set as a CSV file called titanic_original.csv and load it in RStudio into a data frame.

```{r Load Data, echo=TRUE, message=FALSE, warning=FALSE}
# 0: Load the data in RStudio
dfOriginal <- read.csv("titanic_original.csv")
df <- dfOriginal
df
```

The embarked column has some missing values, which are known to correspond to passengers who actually embarked at Southampton. Find the missing values and replace them with S. (Caution: Sometimes a missing value might be read into R as a blank or empty string.)

```{r Port of embarkation, echo=TRUE, message=FALSE, warning=FALSE}
# 1: Port of embarkation
for (i in 1:length(df$embarked)) {
  if(is.na(df$embarked[i]) | is.null(trimws(df$embarked[i])) | trimws(df$embarked[i]) == "" ) {
    df$embarked[i] <- "S"
  }
}
df
```

Calculate the mean of the Age column and use that value to populate the missing values
Why would you pick any of those over the mean (or not)? Mean wont affect many of the statistical calculations.

```{r Age, echo=TRUE, message=FALSE, warning=FALSE}
# 2: Age
# Applying mean is better for missing values as it wont affect any calculations we got to perform on this column
meanAge <- mean(df$age, na.rm = TRUE)
df <- mutate(df, age = ifelse(is.na(age), meanAge, age))
df
```

You’re interested in looking at the distribution of passengers in different lifeboats, but as we know, many passengers did not make it to a boat :-( This means that there are a lot of missing values in the boat column. Fill these empty slots with a dummy value e.g. the string 'None' or 'NA'

```{r Lifeboat, echo=TRUE, message=FALSE, warning=FALSE}
df <- mutate(df, boat = ifelse(is.null(boat) | is.na(boat) | trimws(boat) == "", NA, boat))
df
```

You notice that many passengers don’t have a cabin number associated with them.
* Does it make sense to fill missing cabin numbers with a value?
* What does a missing value here mean?
You have a hunch that the fact that the cabin number is missing might be a useful indicator of survival. 
Create a new column has_cabin_number which has 1 if there is a cabin number, and 0 otherwise.

```{r Cabin, echo=TRUE, message=FALSE, warning=FALSE}
df <- mutate(df, has_cabin_number = ifelse(is.null(cabin) | is.na(cabin) | trimws(cabin) == "", 0, 1))
df
```

Write back to CSV file as titanic_clean.csv
```{r Save output, echo=TRUE, message=FALSE, warning=FALSE}
write.csv(df, file = "titanic_clean.csv")
```